---
title: "BHS_Master_Code"
author: "Operation_Research_BHS_Group"
date: "December 3, 2016"
output: pdf_document
---

Import data:

```{r}
bhs= read.csv(file.choose(), header = TRUE)
```

#Add Specialist Column - Ben Solis-Cohen
```{r}
# All this code is to construct "specs_dict" (see below)
# Specs_dict maps nfh_specialist names to categories like diabetes, diet, and other
bhs2 = bhs
specs = read.csv("NFH Specialists.csv", header = TRUE)
leftout = setdiff(trimws(unique(bhs2$NFH_Resource[bhs2$NFH_Resource != ""])),trimws(specs[,1]))
spec_rownames = c(as.character(specs[,1]),leftout)
specs_dict = c(as.character(specs[,2]),rep("other",length(leftout)))
names(specs_dict) = spec_rownames

# Constructed specs_dict. Now apply to NFH_Resource row to transform names to specialties
bhs2$NFHSpecialist = specs_dict[trimws(bhs2$NFH_Resource)]
bhs = bhs2
```

#Convert ICD-9 to ICD-10 - Ben Solis-Cohen
```{r message=FALSE,eval, FALSE}
### Install pakcages needed for icd conversion.
### Might need to uncomment some lines if you haven't run them already
### Note that you only need to "install" once but need to require/library 
### every time you open a new session
#install.packages("devtools")
library(devtools) 
#devtools::install_github("wtcooper/icdcoder")
library(icdcoder)
require(devtools)
require(dplyr)
require(data.table)
library(icd)
```

```{r,eval = FALSE}
## This code converts all ICD-9 codes to ICD-10 and leaves ICD-10 codes as-is
## Note that the conversion is time-consuming. Don't be surprised if it's taking a long time
for (i in 1:63){
  print(i)
  newDX = matrix(nrow = nrow(bhs),ncol = 1)
  for(j in 1:nrow(bhs2)){ 
    to_convert = bhs[,paste("DX",as.character(i),sep = "")][j]
    icd10_short = as.character(convICD(to_convert, "icd9")[1,2])
    if(is.na(icd10_short)){
      res = as.character(icd_decimal_to_short(to_convert))
    }else{
      res = icd10_short
    }
    newDX[j,1] = res
  }
  bhs[,paste("DX",as.character(i),sep = "")] = newDX
}
```

#Introduce DXiName{1,2,3} vars - Ben Solis-Cohen
```{r,message = FALSE}
library(icd)
```

Below is function to go up the ICD-10 tree using icd10cm2016 table. "s" specifies level to go up. "s" = 6,7,8 are used. See icd10cm2016. "s" is used to index the COLUMNS.
```{r,eval = FALSE}
getA = function (icd10_short,s) {
  res = (as.character(unlist(
    icd10cm2016[which(icd10cm2016$code == icd10_short),][s]
    )))
  if(identical(res,character(0))){
    return ("NA")
  }else{
    return (res)
  }
}
```

```{r,eval=FALSE}
# Introduces DXiName{1,2,3} vars for DX1-DX63 
# Should be using tree compression instead!
bhs2 = bhs
for(i in 1:63){
  print(i)
  #s for specificity
  for(s in 7:8){
    newDXName = matrix(nrow = nrow(bhs),ncol = 1)
    for(j in 1:nrow(bhs2)){
      to_convert = bhs[,paste("DX",as.character(i),sep = "")][j]
      newDX[j,1] = getA(to_convert,s)
    }
    bhs2[,paste("DX",as.character(i),"Name",as.character(s-5),sep = "")] = newDX
  }
}
bhs = bhs2
```

# Add NFH services - Hallee Wong
## Function to record number of specialist encounters for each patient (looking forward from a specific hospital discharge)
```{r}
add_specialists = function (data){ 
    
    #Reorder data by patient ID
    data <- data[order(data$UnitNumber,data$DischargeDateTime),]
    
    #Create zero columns for NFH specialist info
    data$NFH_total <- as.numeric(0)
    data$NFH_nav <- as.numeric(0)
    data$NFH_worker <- as.numeric(0)
    data$NFH_psych <- as.numeric(0)
    data$NFH_bh <- as.numeric(0)
    data$NFH_gmed <- as.numeric(0)
    data$NFH_chf <- as.numeric(0)
    data$NFH_diabetes <- as.numeric(0)
    data$NFH_diet <- as.numeric(0)
    data$NFH_day <- as.numeric(0)
    data$NFH_tobacco <- as.numeric(0)
    data$NFH_other <- as.numeric(0)
    data$keep <- as.numeric(0)
  
    i = 1
    
  while (i <= nrow(data)){
      print(i)
      currentID = trimws(data$UnitNumber[i])
      currentDATE = trimws(data$DischargeDateTime[i])
      data$keep[i] = 1 

      #Find all rows with the same patient ID and discharge date
      j = i
      
      while (trimws(data$UnitNumber[j]) == currentID && 
             trimws(data$DischargeDateTime[j]) == currentDATE && j <= nrow(data)){

        if ((trimws(data$NFH_Resource[j]) != '')){

            if (trimws(data$NFH_Resource[j]) == 'GASPKA'
                ||trimws(data$NFH_Resource[j]) == 'ACCEKR'){
              #print('GASPKA')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_nav[i] <- data$NFH_nav[i] + 1 
              }
              
            else if (trimws(data$NFH_Resource[j]) == 'PIANEL'){
              #print('PIANEL')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_worker[i] <- data$NFH_worker[i] + 1
              }
              
            else if (trimws(data$NFH_Resource[j]) == 'CARTJE'){
              #print('CARTJE')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_psych[i] <- data$NFH_psych[i] + 1
              }
              
            else if (trimws(data$NFH_Resource[j]) == 'UPSOVI'){
              #print('UPSOVI')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_bh[i] <- data$NFH_bh[i] + 1
              }
            
            else if (trimws(data$NFH_Resource[j]) == 'GEYERA'){
              #print('GEYERA')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_gmed[i] <- data$NFH_gmed[i] + 1
              }
              
            else if (trimws(data$NFH_Resource[j]) == 'EMERAL'
                ||trimws(data$NFH_Resource[j]) == 'GILVBR'){
              #print('EMandGI')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_chf[i] <- data$NFH_chf[i] + 1
              }
              
            else if (trimws(data$NFH_Resource[j]) == 'SALZJA'
                ||trimws(data$NFH_Resource[j]) == 'JOLIJI'){
              #print('SAandJo')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_diabetes[i] <- data$NFH_diabetes[i] + 1
              }
              
            else if (trimws(data$NFH_Resource[j]) == 'NELSMA'){
              #print('NELSMA')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_diet[i] <- data$NFH_diet[i] + 1
              }
              
            else if (trimws(data$NFH_Resource[j]) == 'GOLKCA'){
              #print('GOLKCA')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_day[i] <- data$NFH_day[i] + 1
              }
              
            else if (trimws(data$NFH_Resource[j]) == 'AHLIMA'){
              #print('AHLIMA')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_tobacco[i] <- data$NFH_tobacco[i] + 1
              }
            
            else if (trimws(data$NFH_Resource[j]) != ''){
              #print('OTHER')
              data$NFH_total[i] <- data$NFH_total[i] + 1
              data$NFH_other[i] <- data$NFH_other[i] + 1
              }
            
        };
        j = j + 1;
        
      };
      i = j;   
    }
    return(data)
}
```

```{r}
bhs = add_specialists(bhs)
# remove keep == 1 rows (rows that are duplicate discharges)
# These rows only differ by NFH/Education visits and we record
# much of this information in the variables created above.
# However, more could be done in the way of recording *pre-discharge*
# NFH and education visits (as apposed to post-discharge recorded above)
bhs = subset(bhs, keep == 1)
```

##Tree Compression Algorithm - Ben & Inte
```{r}
# Quickly put together. We could think about consolidating into a loop. 
# Algorithm: 
## - Goal: Assure that each disease has a population greater than thresh
## - Method: Group together diseases under the same ancestor in the ICD-10 tree
#### 1) For each ICD-10 code at the most specific level (in the dataset), 
        # If the disease has a population greater than thresh, add to sigDis list
        # Otherwise, add parent to next level list
#### Apply (1) to next level recursively

library(icd)
library(data.table)

SigDis=c()
thresh=40

icdtable=icd10cm2016
row.names(icdtable)=icdtable[,"code"]

level1=table(bhsTree$DX1)
level1[1]=100  # don't want empty strings! Ask Ben/Inte
level2=rep(0,nrow(level1)) #generally size of prev level
names(level2)=rep("none",nrow(level1)) #generally size of prev level

idX=1
for (i in 2:length(level1)){
  count1=level1[i]
  
  if (count1<thresh){
    parent= as.character(icdtable[names(level1)[i],"major"])
    
    if (!(parent %in% names(level2))){ #new parent
      names(level2)[idX]=parent
      level2[parent]=count1
      idX=idX+1
    } else {
      level2[parent]=level2[parent]+count1
    }
  }else{
    disName=icdtable[names(level1)[i],"short_desc"]
    SigDis=c(SigDis,disName)
  }
}

level2=level2[which(names(level2) !="none")]


newtable=as.data.table(icdtable)
icdtable=newtable[,lapply(.SD,function(x){x[1]}),by="major"]
icdtable=as.matrix(icdtable)
row.names(icdtable)=icdtable[,"major"]

level3=rep(0,length(level2)) #generally size of prev level
names(level3)=rep("none",length(level2)) #generally size of prev level


idX=1
for (i in 1:length(level2)){
  count1=level2[i]
  
  if (count1<thresh){
    parent= as.character(icdtable[names(level2)[i],"sub_chapter"])
    
    if (!(parent %in% names(level3))){ #new parent
      names(level3)[idX]=parent
      level3[parent]=count1
      idX=idX+1
    } else {
      level3[parent]=level3[parent]+count1
    }
  }else{
    SigDis=c(SigDis,names(level2)[i])
  }
}

level3=level3[which(names(level3) !="none")]

newtable=as.data.table(icdtable)
icdtable=newtable[,lapply(.SD,function(x){x[1]}),by="sub_chapter"]
icdtable=as.matrix(icdtable)
row.names(icdtable)=icdtable[,"sub_chapter"]

level4=rep(0,length(level3)) #generally size of prev level
names(level4)=rep("none",length(level3)) #generally size of prev level


idX=1
for (i in 1:length(level3)){
  count1=level3[i]
  
  if (count1<thresh){
    parent= as.character(icdtable[names(level3)[i],"chapter"])
    
    if (!(parent %in% names(level4))){ #new parent
      names(level4)[idX]=parent
      level4[parent]=count1
      idX=idX+1
    } else {
      level4[parent]=level4[parent]+count1
    }
  }else{
    SigDis=c(SigDis,names(level3)[i])
  }
}

level4=level4[which(names(level4) !="none")]

other=rep(0,length(level4))
names(other)=rep("none",length(level4))

idX=1
for (i in 1:length(level4)){
  disease=names(level4)[i]
  count=level4[i]
  if (count<20){
    other[idX]=count
    names(other)[idX]=disease
    idX=idX+1
  }else{
    SigDis=c(SigDis,names(level4)[i])
  }
}
other=other[which(names(other) !="none")]  
  
print(SigDis)
print(other)
```

```{r}
# Helper function for tree compression algorithm
# Finds the least common ancestor of the disease with count greater than 40. 
# Requires preprocessing of tree compression algorithm to create sigDis. 
findLCA = function(disName,sigDis,icdtable){
  
  thisRow=icdtable[disName,]
  disName=thisRow[1,"short_desc"]
  
  if (disName %in% sigDis){
    return (disName)
  }
  disName=thisRow[1,"major"]  
  
  if (disName %in% sigDis){
    return (disName)
  }
  disName=thisRow[1,"sub_chapter"]
  
  if (disName %in% sigDis){
    return (disName)
  }
  disName=thisRow[1,"chapter"]
  
  if (disName %in% sigDis){
    return(disName)
  }
  return("Other")
}
```

```{r}
## Wrapper Code to call LCAFinder
## End product is bhs#DX1NameAgg, which is bhs$DX1 where each disease has been
## converted to it's least common ancestor in the ICD-10 tree with population > thresh
icdtable=icd10cm2016
row.names(icdtable)=icdtable[,"code"]
x=apply(as.matrix(bhs$DX1),1,function(disName){return(as.character(findLCA(disName,SigDis,icdtable)))})
bhs$DX1NameAgg=x
```

#Making 'Previous Visits' Variables - Ben Solis-Cohen
For each admission, starting at the last one, iterate up through the other admissions (backwards in time). If a discharge date is within 30 days of the current admission, increment PrevAdmits30. Otherwise, increment PrevAdmitsTot (**which is actually number of previous admissions before 30**). Also, added variable for *same* primary diagnosis. 
```{r}
bhs = read.csv(file.choose())
```

```{r,eval = FALSE}
# This runs with hallee's code already run (keep == 0 columns dropped)
# Copy bhs into bhs2 and sort by admit date
bhs2 = bhs
bhs2 = bhs2[order(bhs$UnitNumber,as.POSIXct(as.character(bhs2$AdmitDateTime), format = "%m/%d/%Y %H:%M")),]
bhs2$NFHSpecialist = trimws(bhs2$NFHSpecialist)

#New blank columns
prevAdmitsTotCol30 = matrix(0,ncol = 1, nrow = nrow(bhs))
prevAdmitsSameCol30 = matrix(0,ncol = 1, nrow = nrow(bhs))
prevAdmitsTotCol = matrix(0,ncol = 1, nrow = nrow(bhs))
prevAdmitsSameCol = matrix(0,ncol = 1, nrow = nrow(bhs))
#Go until second row (first has no previous admits...)
for(curIndex in nrow(bhs2):2){
  print(curIndex)
  curAdmitDate = as.POSIXct(as.character(bhs2$AdmitDateTime[curIndex]), format = "%m/%d/%Y %H:%M")
  curID = bhs2$UnitNumber[curIndex]
  curDisease = bhs2$DX1[curIndex]
  
  checkIndex = curIndex - 1
  checkDischDate = as.POSIXct(as.character(bhs2$DischargeDateTime[checkIndex]), format = "%m/%d/%Y %H:%M")
  checkID = bhs2$UnitNumber[checkIndex]
  checkDisease = bhs2$DX1[checkIndex]
  
  numAdmissions30 = 0
  sameAdmissions30 = 0
  numAdmissions = 0
  sameAdmissions = 0
  while(checkIndex >= 1 && checkID == curID ){

    if(curAdmitDate < checkDischDate){
      print("Should never get here if rows have been dropped")
      checkIndex = checkIndex - 1
      checkID = bhs$UnitNumber[checkIndex]
      checkDisease = bhs2$DX1[checkIndex]
      checkDischDate = as.POSIXct(as.character(bhs2$DischargeDateTime[checkIndex]), format = "%m/%d/%Y %H:%M")
      next
    }

    if(as.numeric(difftime(curAdmitDate,checkDischDate,units = "days")) <= 30){
      numAdmissions30 = numAdmissions30 + 1
      if(checkDisease == curDisease){
        sameAdmissions30 = sameAdmissions30 + 1
      }
    }else{
      numAdmissions = numAdmissions + 1
      if(checkDisease == curDisease){
        sameAdmissions = sameAdmissions + 1
      }
    }
      
    checkIndex = checkIndex - 1
    checkID = bhs$UnitNumber[checkIndex]
    checkDisease = bhs2$DX1[checkIndex]
    checkDischDate = as.POSIXct(as.character(bhs2$DischargeDateTime[checkIndex]), format = "%m/%d/%Y %H:%M")
    
  }
  
  prevAdmitsSameCol30[curIndex] = sameAdmissions30
  prevAdmitsTotCol30[curIndex] = numAdmissions30
  prevAdmitsSameCol[curIndex] = sameAdmissions
  prevAdmitsTotCol[curIndex] = numAdmissions
}

bhs2[,"PrevAdmits30"] = prevAdmitsTotCol30
bhs2[,"PrevAdmitsSame30"] = prevAdmitsSameCol30
bhs2[,"PrevAdmitsTot"] = prevAdmitsTotCol
bhs2[,"PrevAdmitsSameTot"] = prevAdmitsSameCol
View(bhs2)
```

```{r}
bhs = bhs2
```

#Seasonality
```{r}
# Bins an admission into one of periods bins based on AdmitDateTime (this just returns the bin number)
get_season_bin = function(date, periods){
  return(floor(as.POSIXlt(as.character(date), format = "%m/%d/%Y %H:%M")$yday/365*periods))
}
```

```{r}
# Uses get_season_bin to bin admissions in to season bin
# Each split in splits is the number of divisions in the year 
# (4 yields 4 divisions --> 3 month periods and 12 --> monthly divisions)
splits = c(4,6,12)
for(split in splits){
  split_fun = function(date){return(get_season_bin(date,split))}
  bhs[,paste("season",split,sep = "")] = apply(matrix(bhs$AdmitDateTime),1,split_fun)
}
```

#Binary Variable for Readmission within 30/60 days - Intekhab Hossain
```{r}
# Helper function to calculate date difference used in calculation of 30-day readmission variable
dateMinus <- function(later,prev) { 
  as.numeric(difftime(later,prev,units = "days")) 
}
```

```{r}
bhs = bhs[order(bhs$UnitNumber,as.POSIXct(as.character(bhs$AdmitDateTime), format = "%m/%d/%Y %H:%M")),]

bhs$admittedIn30=rep(0,nrow(bhs))
bhs$admittedIn60=rep(0,nrow(bhs))
for (i in 1:(nrow(bhs))){
  print(i)
  thisRow=bhs[i,] 
  thisID=thisRow['UnitNumber']
  thisAdmit=as.POSIXct(as.character(thisRow[['AdmitDateTime']]), format = "%m/%d/%Y %H:%M") 
  thisDischarge=as.POSIXct(as.character(thisRow[['DischargeDateTime']]), format = "%m/%d/%Y %H:%M") 
  for (j in i+1:nrow(bhs)){ 
    testRow=bhs[j,] 
    testID=testRow['UnitNumber']
    testAdmit=as.POSIXct(as.character(testRow[['AdmitDateTime']]), format = "%m/%d/%Y %H:%M") 
    testDischarge=as.POSIXct(as.character(testRow[['DischargeDateTime']]), format = "%m/%d/%Y %H:%M")
    if (testID==thisID){
      if(dateMinus(testAdmit,thisAdmit)>0 & dateMinus(testDischarge,thisDischarge)>0){
        #print(paste(as.Date(thisAdmit)," ",as.Date(thisDischarge)," ",as.Date(testAdmit)," ",as.Date(testDischarge)," ",dateMinus(testAdmit,thisDischarge)))
      
      if(dateMinus(testAdmit,thisDischarge)>0 & dateMinus(testAdmit,thisDischarge)<=30){
        bhs[i,]$admittedIn30=1 
        print("Found a 30-day readmit! YUS!")
        }
      if(dateMinus(testAdmit,thisDischarge)>0 & dateMinus(testAdmit,thisDischarge)<=60){ 
        bhs[i,]$admittedIn60=1
        print("Found a 60-day readmit! YUS!")
        }
      }
    }
    else{
      break
      }
  }
  }
```

# Transform secondary diagnoses to indicator variables
```{r}
# This function gets all secondary diseases with a given specificity (icd-10 level)
# This should be generalized to just use the tree compression algorithm instead of
# hard-coding a specificity (spec) for ALL diseases. 
getAllDiseases = function(df,spec){
    allDiseases = c()
    for(i in 2:63){
      curDiseases = unique(bhs[,paste("DX",i,"Name",spec,sep = "")])
      allDiseases = union(allDiseases,curDiseases)
    }
    return(allDiseases[allDiseases!='NA'])
}
```

```{r}
# Create secondary disease columns
# Each patient gets a 1 if they have that disease and 0 if not
allDiseases = getAllDiseases(bhs,3)
secDiseaseMat = matrix(0,nrow = nrow(bhs),ncol = length(allDiseases))
colnames(secDiseaseMat) = allDiseases

for(i in 1:nrow(bhs)){
  print(i)
  for(j in 2:63){
    if(!is.na(bhs[i,paste("DX",j,"Name3", sep = "")]) && trimws(bhs[i,paste("DX",j,"Name3", sep = "")]) != "NA"){
      secDiseaseMat[i,bhs[i,paste("DX",j,"Name3", sep = "")]] = 1
    }
  }
}
```

# Adding Sessler Indices
```{r}
# Just converts from icd10 to icd9 and appends a D so that the code is 
# readable by icd9cm_sessler_rsi
icd10_to_9 = function(icd10_code){
  return(paste("D",convICD(icd10_code, "icd10")[1,2],sep = ""))
}
```

```{r}
# Create 4 columns "rsi_1yrpod","rsi_30dlos","rsi_30dpod", and "rsi_inhosp"
# These are sessler scores, which correpond to different types of risk. 
# See medicalrisk package documentation package and google sessler index
# for more information

store=matrix(nrow=nrow(bhs),ncol=4)
colnames(store)=c("rsi_1yrpod","rsi_30dlos","rsi_30dpod","rsi_inhosp")

patientDiseases=c()
for (i in 1:63){
  patientDiseases=c(patientDiseases,paste("DX",i,sep=""))  
}

for (i in 1:nrow(bhs)){
  print(i)
  patientDiseases=as.matrix(bhs[i,patientDiseases])
  # remove blanks and NAs
  patientDiseases = patientDiseases[which(sec != '' & !is.na(sec))]
  # Convert to "D" + icd9 version so icd9cm_sessler_rsi can read
  patientDiseases = sapply(patientDiseases,icd10_to_9,USE.NAMES = F)
  if(length(patientDiseases) != 0){
    # append the 4 sessler scores for this patient
    scores=icd9cm_sessler_rsi(patientDiseases)
    store[i,]=as.matrix(scores)
  }
} 
bhs=cbind(bhs,store)
```

## Length of Stay
```{r,eval = FALSE}
# Add length of stay column
bhs$LOS = difftime(as.POSIXct(bhs$DischargeDateTime, format = "%m/%d/%Y %H:%M"),as.POSIXct(bhs$AdmitDateTime,format = "%m/%d/%Y %H:%M"), units = "days")
```

```{r, eval = FALSE}
# Remove columns no longer needed
# Mostly used for variable creation
# We have hopefully already encoded this information by now
# in a more concise way. 
bhs = cbind(bhs,secDiseaseMat)
rem_cols = c()
for(i in 2:63){
  for(j in 1:3){
    rem_cols = c(rem_cols,paste("DX",i,"Name",j,sep=""))
  }
  rem_cols = c(rem_cols,paste("DX",i,sep=""))
  rem_cols = c(rem_cols,paste("DX",i,"Name",sep =""))
}
bhs <- bhs[,!(names(bhs)) %in% rem_cols]
```

```{r}
write.csv(bhs,"final_data_1.csv")
```

